## Load libraries

Firstly, let's load the libraries that we will use.

They will need to be installed first if you don't already have them.

```{r}
#install.packages("readxl")
library(readxl)
#install.packages("RNetCDF")
library(RNetCDF)
```

## Loading and checking the data

Let's look at the data we are working with first:

```{r}
data <- read_excel("meteorological_measurements_time_series_variables.xlsx")
print(data, n=Inf)
```

We have a time series of meteorlogical data; air_temperature, wind_speed and relative_humidity, each recorded at 1 hour intervals. We can ignore that an incorrect date has been added by default. We will see later that we only need the time anyway. 
Most columns headers are taken from the CF standard names. This is a common vocabulary, where parameter names can be searched for online, with descriptions, so that everyone is clear what the data represent.

http://cfconventions.org/standard-names.html

Let's plot the data.

```{r}
plot(data$Time, data$`air_temperature (degC)`, type ="l")
plot(data$Time, data$`wind_speed (m s-1)`, type ="l")
plot(data$Time, data$`relative_humidity (%)`, type ="l")
```

##  Creating NetCDF file

First let's initialise the dataset

```{r}
nc <- create.nc("meteorological_data_time_series.nc")
```
### Defining dimensions

A NetCDF file has dimensions that define the shape of the data. In this case, our only dimension is time. 'time' is a commonly defined dimension name, so to be CF-compliant, we should use that.

```{r}
dim.def.nc(nc, "time", nrow(data))
```

### About time

There are specific recommendations on how time should be stored in netCDF files. I will try to explain briefly here, and there is a nice explanation here too: https://www.unidata.ucar.edu/software/netcdf/time/recs.html

It is most common to have a dimension named "time" as well as a coordinate variable with the same name. Let's discuss the variable first.

The "time" variable has units that count from a user defined origin, for example "hours since 2020-01-01 00:00 UTC" or "days since 2014-01-01". The units may be in years, days, seconds, nanoseconds, etc. Whilst this approach may seem strange at a glance, it allows the times to be stored in conventional numerical formats such as integers, and to our desired precision. This is much more efficient than using a long timestamp string for each coordinate.

And what about the time dimension? Well this defines the shape or grid of the data. In this dataset, we will have 4 variables (time, air_temperature, relative_humidity, wind_speed), and they will all have the same dimension "time". Therefore, the dimension helps the data user to link each of the variables together.

### Defining Variables

A NetCDF has both coordinate variables and data variables. Dimensions define how many grid points there are, coordinate variables define what the actual values are. 

Let's now initialise our variables. We will assign our values later.

```{r}
var.def.nc(nc, "time", "NC_INT", "time")
var.def.nc(nc, "air_temperature", "NC_DOUBLE", "time")
var.def.nc(nc, "wind_speed", "NC_DOUBLE", "time")
var.def.nc(nc, "relative_humidity", "NC_DOUBLE", "time")
```

Now we need a variable attributes for our variables, so that our dataset is compliant with the CF conventions. 

The standard_name should be selected from here: http://cfconventions.org/standard-names.html. Standard names are commonly accepted parameter names with descriptions. By selecting appropriate standard names for your variable, the data user will be clear exactly what the data represent.

The units should match what is provided for the standard name as listed above. You may need to convert your data.

The long_name is more descriptive and can be in your own words.

The coverage_content_type describes what type of data the variable contains

Some help on these variable attributes can be found here: https://commons.esipfed.org/acdd_1-3_references

The data variables should be assigned a fill value. This is a value that is used at grid points where no data exists. We usually choose an unrealistically large value that will obviously appear as a spike to the data user. The fill value is later specified as a variable attribute. Note that coordinate variables should not have fill values as they should not have missing values.

```{r}
att.put.nc(nc, "time", "units", "NC_CHAR", "hours since 2021-02-21T10:00:00Z") # Units for time must be in this format.
att.put.nc(nc, "time", "long_name", "NC_CHAR", "time in hours since 2021-02-21T10:00:00Z")

# It is just coincidence that I have used the same variable name (2nd argument) as the standard name (last argument) here.
att.put.nc(nc, "wind_speed", "standard_name", "NC_CHAR", "wind_speed")
att.put.nc(nc, "wind_speed", "long_name", "NC_CHAR", "Wind speed, magnitude of the 2 horizontal components of velocity, measured at an elevation of 2 meters")
att.put.nc(nc, "wind_speed", "coverage_content_type", "NC_CHAR", "physicalMeasurement")
att.put.nc(nc, "wind_speed", "units", "NC_CHAR", "m s-1")
att.put.nc(nc, "wind_speed", "_FillValue", "NC_DOUBLE", 1e32)

att.put.nc(nc, "air_temperature", "standard_name", "NC_CHAR", "air_temperature")
att.put.nc(nc, "air_temperature", "long_name", "NC_CHAR", "Air temperature in degrees kelvin, measured at an elevation of 2 meters")
att.put.nc(nc, "air_temperature", "coverage_content_type", "NC_CHAR", "physicalMeasurement")
att.put.nc(nc, "air_temperature", "units", "NC_CHAR", "K")
att.put.nc(nc, "air_temperature", "_FillValue", "NC_DOUBLE", 1e32)

att.put.nc(nc, "relative_humidity", "standard_name", "NC_CHAR", "relative_humidity")
att.put.nc(nc, "relative_humidity", "long_name", "NC_CHAR", "Relative humidity of air, the amount of moisture in the air at a certain temperature compared to what the air can hold at that temperature")
att.put.nc(nc, "relative_humidity", "coverage_content_type", "NC_CHAR", "physicalMeasurement")
att.put.nc(nc, "relative_humidity", "units", "NC_CHAR", "1")
att.put.nc(nc, "relative_humidity", "_FillValue", "NC_DOUBLE", 1e32)
```

### Putting variables

Now we need to put the values into that variable

```{r}
var.put.nc(nc, "time", data$Hour)
var.put.nc(nc, "wind_speed", data$`air_temperature (degC)`+ 273.15) # Converting to preferred units defined by the standard_name
var.put.nc(nc, "wind_speed", data$`wind_speed (m s-1)`)
var.put.nc(nc, "wind_speed", data$`relative_humidity (%)`)
```

### Gloabl attributes

Global attributes describe the dataset as a whole. A list of what global attributes must be included can be found here:

https://adc.met.no/node/4

These are based on the ACDD conventions, that you can find here:

https://wiki.esipfed.org/Attribute_Convention_for_Data_Discovery_1-3

We refer to which conventions the file adheres to (including version) in the 'conventions' global attribute.

Additional global attributes can also be included, defined by the user. Make sure that the attribute names you select are understandable. 

To save you from having to write them all out, I have written them in a separate file that you can load in as below.

```{r}
global_attributes <- read_excel("meteorological_measurements_time_series_global_attributes.xlsx")
```

Now we can iterate over each row and write each attribute in turn. We do this by using a special attribute name, NC_GLOBAL. 

```{r}
for (i in 1:nrow(global_attributes)) {
    row <- global_attributes[i,]
    if (row$Format == "NC_DOUBLE") {
      att.put.nc(nc, "NC_GLOBAL", row$AttributeName, "NC_DOUBLE", as.double(row$Content))  
    }
    else {
      att.put.nc(nc, "NC_GLOBAL", row$AttributeName, "NC_CHAR", row$Content)
    }
}
```

We should update some of the attributes to include the current timestamp.

```{r}
dtnow <- format(Sys.time(), tz = "UTC", "%FT%R:%SZ")
att.put.nc(nc, "NC_GLOBAL", "date_created", "NC_CHAR", dtnow)
att.put.nc(nc, "NC_GLOBAL", "history", "NC_CHAR", paste("File created using RNetCDF in R at",dtnow))
```

Finally, we close the file to ensure that it is written and that everything is complete.

```{r}
close.nc(nc)
```

We can then open it up and print it to check if everything looks okay!

```{r}
nc <- open.nc("meteorological_data_time_series.nc")
print.nc(nc)
```
