## Load libraries

Firstly, let's load the libraries that we will use.

They will need to be installed first if you don't already have them.

```{r}
#install.packages('writexl')
library(writexl)
#install.packages("RNetCDF")
library(RNetCDF)
```

## Introducing the data

In this example, we will be loading a depth profile of some Chlorophyll A data. However, this example should be relevant for depth profiles of any data.

Nansen Legacy data can be found via the SIOS data access portal. All Nansen Legacy datasets should be returned when filtering using the 'AeN' collection. Please contact data.nleg@unis.no if you have any problems finding or accessing data.

I have downloaded the following dataset into my directory.

## Loading the data

```{r}
nc <- open.nc("AR_PR_CT_58US_2021710.nc")
print.nc(nc)
```

At a glance, we can see it has 5 dimensions; they they show that there are data from 44 different locations. There are 4363 points of depth. This doesn't mean that there are 4363 samples for every station; more likely there is a lot of 'empty' space in this file where a measurement was not taken at a certain depth. This is neccessary for us to use a single depth dimension for a range of different depth profiles, which each sample different depths.

There is then a whole host of variables and attributes which correspond. The coordinate variables are first, with the same name as their respective dimension. For example TIME(TIME) is the VARIABLE(DIMENSION). The dimension states how many times have been sampled, the variable states what these times are.

Most of the variables have two dimensions; depth and time. Latitude and longitude are only used in coordinate variables, but we can assume here that each coordinate corresponds to a single time. There are other ways to create a netcdf file to more explicitly state this, by having longitude and latitude variables that each have the dimension of time, thus linking them together. An important point to take away is that different people have different ways of doing things, but we should be able to easily understand what has been done and adapt our code accordingly. 

## A closer look

We can access a variable attribute like this:

```{r}
att.get.nc(nc, "PSAL", "units")
```

Or we can use the special "NC_GLOBAL" variable name to access a global attribute, like this:

```{r}
att.get.nc(nc, "NC_GLOBAL", "creator_name")
```

## Writing data to variables

Let's write the data that we are interested in to variables we can easily use later. There is a lot of data here so I am not going to print them!:

```{r}
time <- var.get.nc(nc, "TIME")
print(dim(time))
latitude <- var.get.nc(nc, "LATITUDE")
print(dim(latitude))
longitude <- var.get.nc(nc, "LONGITUDE")
print(dim(longitude))
pressure <- var.get.nc(nc, "PRES")
print(dim(pressure))
psal <- var.get.nc(nc, "PSAL")
print(dim(psal))
temp <- var.get.nc(nc, "TEMP")
print(dim(temp))
var.get.nc(nc, "TIME")
```


Our time variable is currently in 'days since 1950-01-01T00:00:00Z' Let's convert these values to something more useful to us.

```{r}
att.get.nc(nc, "TIME", "units")
date0 <- strptime("01/01/1950 00:00:00", "%m/%d/%Y %H:%M:%S", tz="UTC")
timestamps <- date0 + (3600 * 24 * time)
timestamps
```

## Outputting to Excel

The file doesn't explicitly tell us, but we will assume that each of our 44 coordinates relates to a single one of our 44 times. Before we output these data to Excel, we can first combine them into a dataframe (basically a table).

There are a number of different ways to do this depending on how we want to structure our data. Here, we will create one file for each depth profile. Each file will have multiple columns; pressure, salinity and temperature. We also want to log our latitude, longitude and time information somewhere. In this case, I will dump all this to a separate file, along with a 'profile number' counter. I will use this 'profile number' in each file name for our data. 

WHAT IS WRONG WITH MY TEMPERATURE VARIABLE?

```{r}
for (profile in 1:length(time)) {
  output_filepath_profile <- paste("/home/lukem/ctd_data_profile_",profile,".xlsx", sep="")
  psal_profile <- psal[,profile]
  temp_profile <- temp[,profile]
  pressure_profile <- pressure[,profile]
  df_profile <- data.frame(Pressure = pressure_profile, Temperature = temp_profile, PracticalSalinity = psal_profile)
  write_xlsx(df_profile, output_filepath_profile)
}

df_profiles_overview <- data.frame(Time = timestamps, Latitude = latitude, Longitude = longitude)
write_xlsx(df_profiles_overview, "/home/lukem/ctd_data_profiles_overview.xlsx")
```
