## Load libraries

Firstly, let's load the libraries that we will use.

They will need to be installed first if you don't already have them.

```{r}
#install.packages('writexl')
library(writexl)
#install.packages("RNetCDF")
library(RNetCDF)
```

## Introducing the data

In this example, we will be loading a depth profile of some Chlorophyll A data. However, this example should be relevant for depth profiles of any data.

Nansen Legacy data can be found via the SIOS data access portal. All Nansen Legacy datasets should be returned when filtering using the 'AeN' collection. Please contact data.nleg@unis.no if you have any problems finding or accessing data.

I have downloaded the following dataset into my directory.

## Loading the data

```{r}
data <- open.nc("AeN_SeaWaterChlorophyllA_20191210_P2.nc")
print.nc(data)
```

At a glance, we can see it has 2 dimensions; they denote that there are 168 depth points and 36 characters (NCHAR).

There are 2 coordinate variables, with the same name as the dimensions. 

There are 4 data variables. The CHLOROPHYLL_A, FILTERED_VOLUME and PHAEOPIGMENT variables all have depth as their dimension. This means that the first chlorophyll a value corresponds to the first depth value.

The EVENT_ID has two dimensions, DEPTH and NCHAR. This is therefore a 2 dimensional array. This tells us that there is a maximum of one EVENT_ID for each of the 168 depths, and it has 36 characters - though variables can have missing values, denoted using a fill value. More on this later.

Finally, we have a long list of global attributes.

## A closer look

We can access a variable attribute like this:

```{r}
att.get.nc(data, "CHLOROPHYLL_A", "units")
```

Or we can use the special "NC_GLOBAL" variable name to access a global attribute, like this:

```{r}
att.get.nc(data, "NC_GLOBAL", "creator_name")
```

Let's write our Chlorophyll A data out as a variable we can easily use later:

```{r}
chla <- var.get.nc(data, "CHLOROPHYLL_A")
depth <- var.get.nc(data, "DEPTH")
phaeopigment <- var.get.nc(data, "PHAEOPIGMENT")
filteredvol <- var.get.nc(data, "FILTERED_VOLUME")
plot(chla, -depth)
```

## Outputting to Excel

To output the data to Excel, we can first create a dataframe. 

```{r}
df <- data.frame(Depth = depth, ChlorophyllA = chla, Phaeopigment = phaeopigment, FilteredVolume = filteredvol)
df
```

Let's remove the rows with NAs

```{r}
df <- na.omit(df)
df
```

And then to output

```{r}
write_xlsx(df, '/home/lukem/chlorophyll_a_data_20191210_P2.xlsx')
```